Kelp morphometric properties and the link to hydrodynamic modelling

### 2018-10-03 proposal_draft_2018.Rmd
chapter outline added and references updated. Word and pdf outputs updated. Note that the word out is not great...

---
title: "Kelp morphometric properties and the link to hydrodynamic modelling"
author:
- affiliation: University of the Western Cape
  name: Ross Coppin, AJ Smit, Christo Rautenbach
date: "2018"
output:
  word_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    highlight: zenburn
    toc: yes
  pdf_document:
    toc: yes
    toc_depth: 4
bibliography: mycollection.bib   
---

@@

Wave exposure may be modelled through various methods which range from simple cartographic to more advanced numerical wave models. Cartographical models can be qualitative or quantitative and were designed for the need of wave exposure measures to explain ecological distributions. A simple set of calculations on coastline and wind data, and relatively small input data sets are required. These are regarded as "fetch-based models", which measure the length of open water associated with a site along a straight line. The output of such an approach is a simplified estimate of the potential wave energy for a specific set of sites. Advances in cartographical methods using fetch-based models has allowed for wave exposure measurments for larger areas, and has been suggested as a method for predicting macroalgal community structure [@Burrows2008]. An example of such a model is the "BioEx model" which was developed by @baardseth1970 to estimate wave exposure over large regions. BioEx requires frequency, strength and direction of winds, weighted by degree of exposure within various directions. BioEx is calculated as the sum of the index developed at different spatial scales (local, fjord and open). This method has been used in mapping of marine coastal biodiveristy [@Rinde2004].

@@

A) Determine if the hydrodynamic environment is the main driver of kelp morphology and if this is specific to a location
B) Simulate kelp rafting by means of a hydrodynamic modelling and calibrate this model with *in situ* beach-cast morphometric data.
C) Use the calibrated hydrodynamic model to investigate dispersal of microplastics along the west coast and south-west coast of South Africa.

@@

The hydrodynamic model for coastal flow will be developed using already existing models that have been developed. These models will consist of the SWAN and Delf3D models and the apporach is to use aspects of each model, that are related to coastal flow, from which further development will take place.

Once the model has been developed it will be used to predict kelp morphology at random sites which will then be sampled. The *in situ* kelp morphology data collected at the predicted site will be used to calibrate the model. This process will be repeated until the model can predict coastal flow and kelp morphology within a certian degree of probability. Kelp morphology is driven by regional and local environmental factors, the spatial scales used will be dependent on the spatial patterns of the various abiotic variables being investigated. Fully nested hierarchical design at three spatial scales will be used: regions, locations (within regions) and sites (within locations).

@@

>>>>

---
title: "Kelp morphometric properties and the link to hydrodynamic modelling"
author:
- affiliation: University of the Western Cape
  name: Ross Coppin, AJ Smit, Christo Rautenbach
date: "2018"
output:
  word_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    highlight: zenburn
    toc: yes
  pdf_document:
    toc: yes
    toc_depth: 4
bibliography: LaTeX/mycollection.bib   
---

@@

Wave exposure may be modelled through various methods which range from simple cartographic to more advanced numerical wave models. Cartographical models can be qualitative or quantitative and were designed for the need of wave exposure measures to explain ecological distributions. A simple set of calculations on coastline and wind data, and relatively small input data sets are required. These are regarded as "fetch-based models", which measure the length of open water associated with a site along a straight line. The output of such an approach is a simplified estimate of the potential wave energy for a specific set of sites. Advances in cartographical methods using fetch-based models has allowed for wave exposure measurments for larger areas, and has been suggested as a method for predicting macroalgal community structure [@Burrows2008]. An example of such a model is the "BioEx model" which was developed by @baardseth1970 to estimate wave exposure over large regions. BioEx requires frequency, strength and direction of winds, weighted by degree of exposure within various directions. BioEx is calculated as the sum of the index developed at different spatial scales (local, fjord and open).

@@

A) Determine if the hydrodynamic environment is the main driver of kelp morphology and if this is specific to a location
B) Determine quantitative relationships with kelp morphology and the hydrodynamic environment to create a biological index.
C) Use *in situ* morphological data to calibrate this to a specific area, or ideally site, within a reasonable amount of probability
D) Simulate kelp rafting by means of a hydrodynamic modelling and calibrate this model with *in situ* beach-cast morphometric data.
E) Use the calibrated hydrodynamic model to investigate dispersal of microplastics along the west coast and south-west coast of South Africa.

@@

The hydrodynamic model for coastal flow will be developed using already existing models that have been developed. These models will consist of the SWAN and Delf3D models and the apporach is to use aspects of each model, that are related to coastal flow, from which further development will take place. Once the model has been developed it will be used to predict kelp morphology at random sites which will then be sampled. The *in situ* kelp morphology data collected at the predicted site will be used to calibrate the model. This process will be repeated until the model can predict coastal flow and kelp morphology within a certian degree of probability. Kelp morphology is driven by regional and local environmental factors, the spatial scales used will be dependent on the spatial patterns of the various abiotic variables being investigated. Fully nested hierarchical design at three spatial scales will be used: regions, locations (within regions) and sites (within locations).

@@

## Outline of Chapters

Chapter and sections| Title
-----------------------|------------------------------------------------------------------------
Chapter 1 | Summary
Chapter 1.1| Background
Chapter 1.2| Kelp environmental drivers
Chapter 1.3| Ocean and coastal waves
Chapter 1.4| Hydrodynamic modelling
Chapter 1.5| Aims and objectives
References | Reference list for chapter 
Chapter 2 | Abiotic drivers of morphological adaption of *Ecklonia maxima* along the South African coastline
Chapter 2.1 | Introduction 
Chapter 2.1.2| Waves and kelp morphology 
Chapter 2.1.3| The South African context 
Chapter 2.2 | Aims and objectives
Chapter 2.3 | Methods
Chapter 2.4 | Results
Chapter 2.5 | Discussion
References | Reference list for chapter
Chapter 3 | Using kelp morphology as an index for calibrating coastal-flow model
Chapter 3.1 | Introduction
Chapter 3.1.2 | The SWAN model design and application
Chapter 3.1.3 | SWAN as a platform for further development
Chapter 3.2 | Aims and objectives
Chapter 3.3 | Methods
Chapter 3.4 | Results
Chapter 3.5 | Discussion
References | Reference list for chapter
Chapter 4 | Application of modified coastal-flow model
Chapter 4.1 | Introduction
Chapter 4.1.2 | Plastic plloution in the marine environment
Chapter 4.1.3 | Application of hydrodynamic modelling in dispersal of plastics
Chapter 4.1.4 | Plastic pollution in SOuth Africa
Chapter 4.2 | Aims and objectives
Chapter 4.3 | Methods
Chapter 4.4 | Results
Chapter 4.5 | Discussion
References | Reference list for chapter
Chapter 5 | Synthesis
Chapter 5.1 | Synthesis of Chapters
Chapter 5.2 | Challenges and limitations
Chapter 5.3 | Gaps and future research
Chapter 5.4 | Concluding remarks
